---
title: "Visualisation - Lab 4 - Groups A7"
author: "Malte Grönemann and Varshith Konda"
date: "26/09/2020"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

library(plotly)
library(ggplot2)
library(seriation)
library(dplyr)
library(lattice)
```

# Assignment 1

The file prices-and-earnings.txt shows a UBS report
comparing prices, wages, and other economic conditions in cities around the world. Some of
the variables measured in 73 cities are Cost of Living, Food Costs, Average Hourly Wage,
average number of Working Hours per Year, average number of Vacation Days, hours of work
(at the average wage) needed to buy an iPhone, minutes of work needed to buy a Big Mac,
and Women’s Clothing Cost.

## 1.1

Import the data to R and keep only the columns with the following numbers:
1,2,5,6,7,9,10,16,17,18,19. Use the first column as labels in further analysis.

```{r data1}
prices_and_earnings <- read.delim("prices-and-earnings.txt", row.names = 1, stringsAsFactors = FALSE)
prices_and_earnings <- prices_and_earnings[c(1, 4:6, 8, 9, 15:18)] 
# mentioned variables - 1 because of using 1 as rownames and not importing it as a variable
colnames(prices_and_earnings) <- c("Food_Costs", "iPhone", "Clothing", "Hours", "Wage", "Vacation", "BigMac", "Bread", "Rice", "Goods_Services")

pae_scaled <- scale(prices_and_earnings)
```

## 1.2 

Plot a heatmap of the data without doing any reordering. Is it possible to see clusters,
outliers?

```{r heatmap_unsorted}
plot_ly(z = pae_scaled,
        x = colnames(pae_scaled),
        y = rownames(pae_scaled),
        type = "heatmap") %>%
  layout(title = "Unsorted Heatmap")
```

In the unsorted heatmap, seeing relations between variables or similarities between groups of cities is not possible. Basically, just outliers are identifiable by their color like that bread prize is very high in Manila and the number of vacation days is very low in Mexico City.

## 1.3

Compute distance matrices by a) using Euclidian distance and b) as one minus correlation.
For both cases, compute orders that optimize Hamiltonian Path Length and use Hierarchical
Clustering (HC) as the optimization algorithm. Plot two respective heatmaps and state which
plot seems to be easier to analyse and why. Make a detailed analysis of the plot based on
Euclidian distance.

```{r heatmaps_sorted}
d_euc_cols <- dist(t(pae_scaled))
d_euc_rows <- dist(pae_scaled)
order_euc_cols <- get_order(seriate(d_euc_cols, method = "GW"))
order_euc_rows <- get_order(seriate(d_euc_rows, method = "GW"))
pae_euc <- pae_scaled[rev(order_euc_rows), order_euc_cols]

plot_ly(z = pae_euc,
        x = colnames(pae_euc),
        y = rownames(pae_euc),
        type = "heatmap") %>%
  layout(title = "Ordered Heatmap using the Euclidian Distance")

d_cor_cols <- as.dist(1 - cor(pae_scaled))
d_cor_rows <- as.dist(1 - cor(t(pae_scaled)))
order_cor_cols <- get_order(seriate(d_cor_cols, method = "GW"))
order_cor_rows <- get_order(seriate(d_cor_rows, method = "GW"))
pae_cor <- pae_scaled[rev(order_cor_rows), order_cor_cols]

plot_ly(z = pae_cor,
        x = colnames(pae_cor),
        y = rownames(pae_cor),
        type = "heatmap") %>%
  layout(title = "Ordered Heatmap using the Pearson Distance")
```

Both procedures result in the same ordering of the variables but the ordering of the cities varies. The heatmap using the euclidian distance is easier to interpret because cities that have high absolute values in most considered variables are located at the fringes while cities that are average in everything are located in the middle.

Generally, low amounts of time one has to work for the individual consumer goods are of course interrelated and coincide with high wages, fewer working hours, many vacation days and high overall standards. The cities exhibiting this pattern are typically located in Western Europe and North America. The cities that are relatively average on all dimension tend to be in Eastern Europe. The opposite pattern with lower wages, less goods and services, fewer vacation days and longer working hours as well as higher amounts of time needed to buy certain goods despite lower absolute costs are geographically located in South America and especially Africa.

## 1.4

Compute a permutation that optimizes Hamiltonian Path Length but uses Traveling Salesman
Problem (TSP) as solver. Compare the heatmap given by this reordering with the heatmap
produced by the HC solver in the previous step – which one seems to be better? Compare
also objective function values such as Hamiltonian Path length and Gradient measure
achieved by row permutations of TSP and HC solvers (Hint: use criterion() function)

```{r TSP}
order_TSP_cols <- get_order(seriate(d_euc_cols, method = "TSP"))
order_TSP_rows <- get_order(seriate(d_euc_rows, method = "TSP"))
pae_TSP <- pae_scaled[rev(order_TSP_rows), order_TSP_cols]

plot_ly(z = pae_TSP,
        x = colnames(pae_TSP),
        y = rownames(pae_TSP),
        type = "heatmap") %>%
  layout(title = "Heatmap using the Euclidian Distance and TSP")

criterion(d_euc_rows)
```

With the TSP optimization, the advantage describe above of the euclidian distance based heatmap vanishes. The patterns are less clear and the typical geographic clustering cannot be seen as clearly anymore.

## 1.5

Use Ploty to create parallel coordinate plots from unsorted data and try to permute the
variables in the plot manually to achieve a better clustering picture. After you are ready with this, brush clusters by different colors and comment about the properties of the clusters:
which variables are important to define these clusters and what values of these variables are
specific to each cluster. Can these clusters be interpreted? Find the most prominent outlier
and interpret it.

```{r parallelcoordinates}
plot_ly(type = "parcoords",
        line = list(color = ~prices_and_earnings$Wage),
        dimensions = list(
          list(values = ~prices_and_earnings$iPhone,
               range = range(prices_and_earnings$iPhone),
               label = "iPhone"),
          list(values = ~prices_and_earnings$Hours,
               range = range(prices_and_earnings$Hours),
               label = "Hours"),
          list(values = ~prices_and_earnings$Vacation,
                 range = range(prices_and_earnings$Vacation),
               label = "Vacation"),
          list(values = ~prices_and_earnings$Goods_Services,
               range = range(prices_and_earnings$Goods_Services),
               label = "Goods_Services"),
          list(values = ~prices_and_earnings$Clothing,
               range = range(prices_and_earnings$Clothing),
               label = "Clothing"),
          list(values = ~prices_and_earnings$Wage,
               range = range(prices_and_earnings$Wage),
               label = "Wage"),
          list(values = ~prices_and_earnings$Food_Costs, 
               range = range(prices_and_earnings$Food_Costs),
               label = "Food"),
          list(values = ~prices_and_earnings$Bread,
               range = range(prices_and_earnings$Bread),
               label = "Bread"),
          list(values = ~prices_and_earnings$BigMac,
               range = range(prices_and_earnings$BigMac),
               label = "BigMac"),
          list(values = ~prices_and_earnings$Rice,
               range = range(prices_and_earnings$Rice),
               label = "Rice")
        )) %>%
  layout(title = "Parallel Coordinates Plot")
```

I would suggest three clusters, similar to the ones discussed for the heatmaps: cities with high living standards (high clothing index, few working hours, lots of vacation) and low time to work to be able to buy consumer goods. The second cluster is less extreme and is typically in the middle of the distributions or follows the trajectories of the first cluster but less extremely. The third one is the opposite of cluster 1. Unsurprisingly, they can largely identified by the mean net wage in each city. The higher the wages, the fewer one has to work for a given standard and given the relation between high wages and higher welfare expenses and living standards especially in Europe, these relations can largely be explained. This is higlited in the plot by assignig the colour of the line to the wage.

## 1.6
 
Use the data obtained by using the HC solver and create a radar chart diagram with
juxtaposed radars. Identify two smaller clusters in your data (choose yourself which ones)
and the most distinct outlier.

```{r radar}
Ps <- list()

for (i in 1:nrow(pae_euc)){
  Ps[[i]] <- htmltools::tags$div(
    plot_ly(type = 'scatterpolar', 
            r = as.numeric(pae_euc[i,-1]),
            theta = colnames(pae_euc)[-1], 
            fill = "toself") %>%
      layout(title = rownames(pae_euc)[i]), 
            style = "width: 25%;")
}

h <-htmltools::tags$div(style = "display: flex; flex-wrap: wrap", Ps)

htmltools::browsable(h)
```

To the second cluster of primarily average values belong also some high-profile cities in Asia like Shanghai, Beijing, Bangkog and Taipeh.

In some Middle Eastern cities, we can observe average values in most variables but both long working hours and many vacation days.

Japanese and Oceanic cities typically belong to the first cluster from the heatmap together with Western Europe and Noth America.

An interesting outlier is Hong Kong with very long working hours, high food costs but low values on most other variables. This probably relates to the special status of Hong Kong.

## 1.7

Which of the tools you have used in this assignment (heatmaps, parallel coordinates or radar
charts) was best in analyzing these data? From which perspective? (e.g. efficiency, simplicity,
etc.)

In my mind, most clusters, their geographic distribution and outlier analysis was already feasible with the heatmap using euclidian distance and HC. At the same time, the heatmaps are most efficient in their use of space and relatively simple to understand. The parallel coordinate plot has the advantage of showing the amount of variation in absolute values in the respective variables. And since it does not need scaling, values are more intuitive in that plot. The radar plots are suited for analysis of each individual case but the easy to identify large patterns is lost. This method is also very ineffiecient in the use of space.




# Assignment 2

```{r , warning=FALSE}
adult_data <- read.csv("adult.csv",header = FALSE)
names(adult_data) <- c("age","workclass","fnlwgt","education","education_num","marital_status","occupation","relationship","race","sex","capital_gain","capital_loss","hours_per_week","native_country","income_level")
```

## Question 1 

Use ggplot2 to make a scatter plot of Hours per Week versus age where observations are
colored by Income level. Why it is problematic to analyze this plot? Make a trellis plot of the same kind where you condition on Income Level. What new conclusions can you make here? 
```{r question 1, warning=FALSE}
plot1 <- ggplot(data  = adult_data,aes(x= hours_per_week , y = age , color =income_level))+
   labs(title = "Plot 2.1 Scatterplot of weekly Hours by Age",
       x = "weekly work hours",
       y = "age",
       colour = "Income Level")+
  geom_point()
  
plot1

plot2 <- qplot(hours_per_week,age, data=adult_data, facets=~income_level)+
  labs(title = "Plot 2.2 Trellis plot of weekly Hours by Age separated by Income Level",
       x = "weekly work hours",
       y = "age")

plot2


```


As we plotted both income level ranges in plot 2.1 we observe over plotting due in the plot,but we can interpret that high number of people earning less than or equal to fifty thousand is higher.

In plot 2.2 using trellis plot we have divided the scateer plot into two plots according to their income levels,in this plot using perception we can depict that peaople earning less than fifty thousand is higher,people  weekly hours  mostly lies between zero and 75,

Income level is higher than fifty thousand in population working 25 to 55 hours weekly.


## Question 2

Use ggplot2 to create a density plot of age grouped by the Income level. Create a trellis plot
of the same kind where you condition on Marital Status. Analyze these two plots and make
conclusions.
```{r question 2 , warning =FALSE}
plot3 <- 
  ggplot(adult_data , aes(age,fill = income_level))+
  geom_density(alpha = 0.50)+
  labs(title  = "2.3 density plot of Ages grouped by income level")

plot3


plot4 <-  ggplot(adult_data , aes(age,fill = adult_data$income_level))+
  geom_density(alpha = 0.50)+
  facet_wrap(~adult_data$`marital_status`)+
  labs(title = "2.4 Trellis plot of income and age whith marital status")
  
plot4
```

from plot 2.3 we can interpret that population earning less than fifty thousand are higher from ages 10 to 25 gradual decrease was observed.

Population with income level higher were higher number between ages thirties to late forties there was a sharp descent was observed.

In plot 2.4 it is clearly observed that almost all the groups have similar densities on income levels but there were spikes in densities were found in never-married and married-spouse-absent.


The percentage of population who are never married earned more in their early twenties.

Married-AF-spouses were more in from early twenties to early fifties.

Widowed population in  the both income groups are similar and are in more percentage from the ages 50 to 75.



## Question 3 

Filter out all observations having Capital loss equal to zero. For the remaining data, use Plotly
to create a 3D-scatter plot of Education-num vs Age vs Captial Loss. Why is it difficult to
analyze this plot? Create a trellis plot with 6 panels in ggplot2 in which each panel shows a
raster-type 2d-density plot of Capital Loss versus Education-num conditioned on values of
Age (use cut_number() ) . Analyze this plot.

```{r question 3 , warninig = FALSE}

CL_data <-  filter(adult_data ,capital_loss > 0)

plot5 <-  plot_ly(CL_data,x = ~education_num  , y = ~age,z = ~capital_loss)%>%
          add_markers()%>%
          labs(title = "2.5 3D scatter plot Education number vs age vs capital loss ")

 

plot5


plot6 <- ggplot(CL_data, aes(x = capital_loss , y = education_num ))+
  stat_density_2d(geom = "raster",
                  aes(fill = after_stat(density)),
                  contour = FALSE)+
          facet_wrap(~cut_interval(age,n = 6))+
  labs(title = "2.6 raster plot of Education number vs capital loss" )


plot6
```

In plot 2.5 we have created an 3d scatter plot for education number , age with Z axis of capital loss.

higher density of points were observed between 1000 to 2500 of capital loss
With this kind of plot there was lot of over plotting done, and we can't practically analyze lot more information from this plot.

The information presented by 2.6 raster plot   to us is also very minimal.

As discussed in 2.5 plot the high density of capitol loss was found majorly from 1000 to 2500 range.



Education number and capitol loss was higher from ages late teens to early sixties.


highest age density was found in 53.5 to 65.7 as observed in pane 4.

There was very low density observed in from  from the ages 77.8 to 90 is. 








## Question 4

Make a trellis plot containing 4 panels where each panel should show a scatter plot of Capital
Loss versus Education-num conditioned on the values of Age by a) using cut_number() b)
using Shingles with 10% overlap. Which advantages and disadvantages you see in using
Shingles?

```{r question 4 }
plot7 <- ggplot(CL_data) +
  aes(x =  capital_loss,
      y = education_num) +
  geom_point() +
  facet_wrap(~cut_number(CL_data$age, 10)) +
  labs(title = "2.7 Scatter Plots of Education and Capital Loss by Age Groups",
       x = "capital loss",
       y = "education number")
plot7

shin <- equal.count(CL_data$age,number = 10 ,overlap = 0.1)


plot8 <- ggplot(CL_data) +
          aes(x = capital_loss,
              y = education_num) +
          geom_point() +
          facet_wrap(shin)+
          labs(title = "2.8  Scatter Plots of Education and Capital Loss by Age Groups",
          x = "capital loss",
          y = "Education num ")
  
plot8
```


from plot 2.7 we observed Education number and education number  for the population mostly lie between 8,16 and 1000 to 2500  especially from the age of  26 and above.


 
The plot 2.8  with intervals of 10& shingles present us with  more insight into data as we get more detailed data points for every age in the data set.

There is a problem of over plotting observed.

2.7 has advantage of less over plotting and more insight into different  age groups.


# Contributions

The main responsibility for Assignment 1 obliged to Malte Grönemann while for Assignment 2 it was the responsibility of konda varshith.

