---
title: "lab - 2 assignment 1"
author: "konda varshith , malte groenemann"
date: "12/09/2020"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(plotly)
```

## Introduction

The data contains the information about different contents of olive oils hailing from different regions in Italy ,Describing about different Regions , areas and different acids present in them.


##  Loading the data 

```{r loading the data }

    olive <- read.csv("olive.csv")

```



## Question 1

```{r 1.1,warning=FALSE}
graph1 <- ggplot(data = olive,aes( x= palmitic, 
                                   y= oleic,
                                   color=linoleic))+
          geom_point()+
          labs(title ="Dependency of palmitic on oleic",x="palmitic",y="oleic")
          graph1

```
The graph depicts the dependence of palmitic on Oleic where colored by Linoleic.

```{r 1.2 ,warning=FALSE}
    

    Linoleic_interval_data <- cut_interval(olive$linoleic,4)
    
    graph2 <- ggplot(data = olive,aes( x= palmitic, 
                                       y= oleic,
                                       color=Linoleic_interval_data))+
    geom_point()
    
    graph2

```



## Question 2 

```{r 2.1,  warning=FALSE}


    graph3 <- ggplot(data = olive,aes( x= palmitic,
                                       y= oleic,
                                       color=Linoleic_interval_data))+
              geom_point()
              
              graph3

graph4 <-  ggplot(data =olive,aes(x=palmitic,
                                  y=oleic,
                                  size=Linoleic_interval_data))+
          geom_point()
           graph4


graph5 <-  ggplot(data = olive,aes(x=palmitic,
                                   y=oleic,
                                   geom_spoke(Linoleic_interval_data)))+
    geom_point()
graph5

```



## Question 3 

```{r 2.3,warning = FALSE}
graph6<- ggplot(data = olive,aes(x= oleic ,y= eicosenoic,
                                 color = Region))+
  geom_point()
graph6
```

#converting numeric data into categorical data 

```{r }
region_categorial <- cut(olive$Region, breaks = c(0,1,2,3),labels = c("North","South","Sardinia islands"))

graph7 <-  ggplot(data = olive,aes(x= oleic ,
                                   y=eicosenoic,
                                   color = region_categorial))+
            geom_point()
            graph7

```


##Question4 

```{r}
linoleic_class_data <- cut_interval(olive$linoleic,3)
palmitic_class_data <- cut_interval(olive$palmitic,3)
palmitoleic_class_data <- cut_interval(olive$palmitoleic,3)
graph8 <-  ggplot(data = olive ,aes(x=oleic,y=eicosenoic,
                                    color = linoleic_class_data,
                                    shape =palmitic_class_data,
                                    size = as.numeric(palmitoleic_class_data) ))+
  geom_point()
graph8

```
##Question 5 


````{r question 5}

graph9 <-  ggplot(data = olive,aes(x=oleic ,y= eicosenoic,color =Region,
                                   shape = palmitic_class_data,
                                   size = as.numeric(palmitoleic_class_data)))+
  geom_point()
graph9
```



##Question 6 
```{r question 6}
graph10 <- plot_ly(data = olive , type='pie',labels = ~olive$Area,values = olive,showlegend=FALSE,showlabels=FALSE)
graph10

```




##question 7 
```{r question 7 }
graph11 <- ggplot(data = olive,aes(x=olive$linoleic,y= olive$eicosenoic))+
  geom_point()+
    geom_density_2d()
graph11
```