---
title: "Visualisation - Lab 2 - Group A7"
author: "Malte Grönemann and Varshith Konda"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(plotly)

olive <- read.csv("olive.csv")
```

# Assignment 1

## Introduction

The data contains the information about different contents of olive oils hailing from different regions in Italy, describing about different regions, areas and different acids present in them.

## 1.1

Create a scatterplot in Ggplot2 that shows dependence of Palmitic on Oleic in which
observations are colored by Linoleic. Create also a similar scatter plot in which you divide
Linoleic variable into fours classes (use cut_interval() ) and map the discretized variable to
color instead. How easy/difficult is it to analyze each of these plots? What kind of perception
problem is demonstrated by this experiment?



```{r 1.1, warning=FALSE}
graph1 <- ggplot(data = olive,
                 aes(x = palmitic,
                     y= oleic,
                     color=linoleic)) +
                geom_point() +
                labs(title = "1.1 | Dependency of palmitic on oleic",
                     x = "palmitic",
                     y = "oleic")
graph1
```



```{r 1.2, warning=FALSE}
    
Linoleic_interval_data <- cut_interval(olive$linoleic,4)
    
graph2 <- ggplot(data = olive,
                 aes(x = palmitic, 
                     y = oleic,
                     color = Linoleic_interval_data)) +
                  geom_point()+
  labs(title = "1.2 | Dependency of palmitic on oleic ",
                     x = "palmitic",
                     y = "oleic")
graph2
```


In the plot 1.1 the points on the plot are colored by the linoleic variable,and in plot 1.2 linoleic variable was divided into four classes to color the plot.

The only channel capacity feature in both plots is hues,it is difficult for human vision to distinguish between the hue variances of the points with a pre-attentive mechanism in plot 1.1, even with a attentive mechanism, it is not possible to differentiate between the points, and there could be a risk of wrong decision due to over plotting the points.

In plot 1.2 , humans can preattentively identify the bright hue in plot, despite of overplotting the points can be differentiated easily.


## 1.2

Create scatterplots of Palmitic vs Oleic in which you map the discretized Linoleic with four
classes to:
a. Color
b. Size
c. Orientation angle (use geom_spoke() )
State in which plots it is more difficult to differentiate between the categories and
connect your findings to perception metrics (i.e. how many bits can be decoded by a
specific aesthetics

```{r 2.1, warning=FALSE}
graph3 <- ggplot(data = olive,
                 aes(x = palmitic,
                     y = oleic,
                     color = Linoleic_interval_data)) +
              geom_point()+
              labs(title= "1.3 | Palmitic and Oleic coloured by discretized lenoleic",
                   x="palmitic",
                   y="oleic")
graph3

graph4 <-  ggplot(data = olive,
                  aes(x = palmitic,
                      y = oleic,
                      size = Linoleic_interval_data)) +
                  geom_point()+
                   labs(title= "1.4 | Palmitic and Oleic size by discretized lenoleic",
                   x="palmitic",
                   y="oleic")
graph4

olive$linoleic_angle <- as.numeric(cut_interval(olive$linoleic,4,labels = c(0,.5*pi,1*pi, 2*pi)))


graph5 <-  ggplot(data = olive,
                  aes(x = palmitic,
                      y = oleic,
                      angle = linoleic_angle,
                      radius  = 50))+
          geom_point()+
                   labs(title= "1.5 | Palmitic and Oleic orientation angle by discretized lenoleic",
                   x="palmitic",
                   y="oleic")+
          geom_spoke()
graph5
```

George Miller (Ward, Grinstein and Keim 2015, 127-9) measured the levels of channel capacities in bits for each hues, shapes and lengths, channel capacities are measured in levels or  bits.

Using the George Miller's experimentation, in plot 1.3 the points on the graph can be distinguished by different level of hues. Hence , 3.1 bits can be encoded for this aesthetic.

According to channel capacities for the *size of squares*, humans were able to classify only just 4 to 5 levels, this suggests that it is difficult for humans to identify sizes correctly. Although plot 1.4 has four and therefore an acceptable number of levels, overplotting of points can be  clearly observed and the use of single hue for points can lead to wrong perception. But even though we have 4 levels, the 2.2 bits that can be encoded for this aesthetic are very low.
 
Experiments for the channel capacities for *line geometries* (Ward, Grinstein and Keim 2015, 127-9) suggest, humans are able to classify only just 2.8 - 3.3 bits in this aesthetic. This suggests that it is difficult for humans to identify orientation angle, In plot 1.5 this can be clearly noticed, 3 bits can be encoded in plot 1.5. 

Among the three plots , plot 1.4 is hardest to differentiate between the categories, mainly due to overplotting but also a low channel capacity. Best is the plot using distinct hues.



## 1.3
Create a scatterplot of Oleic vs Eicosenoic in which color is defined by numeric values of
Region. What is wrong with such a plot? Now create a similar kind of plot in which Region is a
categorical variable. How quickly can you identify decision boundaries? Does preattentive or
attentive mechanism make it possible?

```{r 3.1, warning = FALSE}
graph6 <- ggplot(data = olive,
                 aes(x = oleic,
                     y = eicosenoic,
                     color = Region)) +
                     geom_point()+
          labs(title = "1.6 | Scatterplot of Oleic and Eicosenoic",
               x = "oleic",
               y= "eicosenoic")
graph6
```

## Converting numerical data into categorical data 

```{r 3.2 conversion}
region_categorial <- cut(olive$Region, 
                         breaks = c(0,1,2,3),
                         labels = c("North","South","Sardinia islands"))

graph7 <- ggplot(data = olive,
                 aes(x = oleic,
                     y = eicosenoic,
                     color = region_categorial)) +
              labs(title = "1.7 | Scatterplot of Oleic and Eicosenoic",
               x = "oleic",
               y= "eicosenoic")+
                  geom_point()
graph7
```


In both plots, the only channel capacity features are hues and brightness. It is difficult for human vision to distinguish between the hue variances of the points in plot 1.6 with a pre-attention mechanism, even with an attentive mechanism, it is not possible to differentiate between the points. Overplotting of points is observed for two categories of points, this may lead to slow identification of the boundaries.

Boundaries between two  regions overlapping cannot be determined with pre-attentively.

Decision boundaries can be easily established in 1.7, as the boundary between two groups of elements of the same visual feature can be identified pre-attentively.

## 1.4

Create a scatterplot of Oleic vs Eicosenoic in which color is defined by a discretized Linoleic (3
classes), shape is defined by a discretized Palmitic (3 classes) and size is defined by a
discretized Palmitoleic (3 classes). How difficult is it to differentiate between 27 = 3 * 3 * 3
different types of observations? What kind of perception problem is demonstrated by this
graph?

```{r 4.1}
linoleic_class_data <- cut_interval(olive$linoleic, 3,labels = c("low","medium","high"))
palmitic_class_data <- cut_interval(olive$palmitic, 3,labels = c("low","medium","high"))
palmitoleic_class_data <-as.numeric( cut_interval(olive$palmitoleic, 3,labels = c("low","medium","high")))

graph8 <- ggplot(data = olive,
                 aes(x = oleic,
                     y = eicosenoic,
                     color = linoleic_class_data,
                     shape = palmitic_class_data,
                     size = palmitoleic_class_data)) +
              labs(title ="1.8 | oleic vs eicosenoic",
                   x = "oleic",
                   y= "eicosenoic")+
                  geom_point()
graph8


```


In the plot 1.8 , the points are categorized using three different features "color" ,"shape" and "size" using "linoleic" , "palmitic" and "palmitoleic" variables respectively. 
Use of various channel capacites together is also known as *Multi-dimensional stimuli*.

Even when using multiple features, a maximum of 7.2 bits were encoded, adding multiple stimuli reduced the discernibility of individual features (Ward, Grinstein and Keim 2015, 127-9).

Along with channel capacities, are are a few other issues found while using multiple aesthetics. The plot has overplotting which in turn make it difficult to see overlapping data points and it has too many classes to understand. This might result in troubled interpretations.

## 1.5

````{r question5}
graph9 <- ggplot(data = olive,
                 aes(x=oleic,
                     y= eicosenoic,
                     color =Region,
                     shape = palmitic_class_data,
                     size = as.numeric(palmitoleic_class_data))) +
  labs(title ="1.9 | oleic  vs eicosenoic",
        x = "oleic",
         y = "ecosenoic")+
                  geom_point()
graph9
``` 


In plot 1.9, boundaries can be found to a limited extent. We can observe a boundary between regions but the two regions present at the bottom are worse to tell apart. 

According to Treisman's theory of pre-attentive processing, conjuction of features requires serial search between the maps.

### Layers of Observations using Treisman's Mapping

We can observe the boundaries between three regions by three different hues.

In next layer, we can distinguish the shape of the points, two regions at the bottom have a higher number of triangles, the central cluster consists of squares and triangles. But due to overplotting, much specifications cannot be determined.

Few ircles were found as outliers on the plot.


## 1.6

Use Plotly to create a pie chart that shows the proportions of oils coming from different
Areas. Hide labels in this plot and keep only hover-on labels. Which problem is demonstrated
by this graph?
 
```{r question 6, warning=FALSE}
graph10 <- plot_ly(data = olive, 
                   type='pie',
                   labels = ~Area,
                   values = olive,
                   showlegend = FALSE,
                   showlabels = FALSE,
                   textinfo = 'none'
                   )

graph10
```

As there are no labels on the pie chart, it is hard to understand the representation of the colours the plot. As we cannot identify the angle differences without labels, this pie chart can mislead portions of the chart, these plots are not ideal for static reports.

Good principles of visualization suggest that a plot should possess a proper legend at 100 percent scale and all the necessary information to interpret it.

## 1.7

Create a 2d-density contour plot with Ggplot2 in which you show dependence of Linoleic vs
Eicosenoic. Compare the graph to the scatterplot using the same variables and comment why
this contour plot can be misleading.

```{r question7}
graph11 <- ggplot(data = olive,
                  aes(x = linoleic,
                      y = eicosenoic)) +
                  geom_density_2d()+
                  labs(title ="1.11 | 2d density plot of lenoleic and eicosenoic")




graph12 <- ggplot(data = olive,
                  aes(x = linoleic,
                      y = eicosenoic)) +
                  geom_point() +
       
             labs(title ="1.12 | Scatter plot of lenoleic and eicosenoic")
graph11
graph12
```

Contour lines do not define the specific relationship of the data, this mapping technique is not useful to know the density of the variable.
If we look at the scatterplot, there are some points between 800 and 1000, but the 2d density reveals that the region is very empty.
In the density plot, the plot did not define the outliers of the given data, without the outlier information some key information might be missed.


# Assignment 2

## 2.1)

The data set baseball-2016.xlsx contains information about the scores of baseball teams in USA in 2016, such as: Games won, Games Lost, Runs peer game, At bats, Runs, Hits, Doubles, Triples, Home runs, Runs batted in, Bases stolen, Time caught stealing, Bases on Balls, Strikeouts, Hits/At Bats, On Base Percentage, Slugging percentage, On base+Slugging, Total bases, Double plays grounded into, Times hit by pitch, Sacrifice hits, Sacrifice flies, Intentional base on balls, and Runners Left On Base.

Load the file to R and answer whether it is reasonable to scale these data in order to perform
a multidimensional scaling (MDS).

```{r data2}
library(readxl)
baseball <- read_excel("baseball-2016.xlsx")
baseball_2 <- baseball[ , -(1:2)] # removing team name and league
baseball_2 <- scale(baseball_2)

anyNA(baseball_2)
```

The data is generally appropriate for multidimensional scaling (MDS) but some variables are count data. Therefore it is more appropriate to focus on order than distance which can be achieved by using non-metric MDS. Before doing the procedure, we want to exclude the variables defining the teams, the team name and the league they play in. Fortunately, the dataset contains no missing values that could mess with the procedure.

## 2.2)

Write an R code that performs a non-metric MDS with Minkowski distance=2 of the data (numerical columns) into two dimensions. Visualize the resulting observations in Plotly as a scatter plot in which observations are colored by League. Does it seem to exist a difference between the leagues according to the plot? Which of the MDS components seem to provide the best differentiation between the Leagues? Which baseball teams seem to be outliers?

```{r nmds, message=FALSE, warning=FALSE}
library(MASS)
library(plotly)

d <- dist(baseball_2, 
          method = "minkowski", 
          p = 2) # power of the minkowski distance set to 2

fit <- isoMDS(d, k = 2) # fitting a nmds for k = 2 dimensions

baseball$x <- fit$points[ , 1]
baseball$y <- fit$points[ , 2]

plot_ly(data = baseball,
       x = ~x,
       y = ~y,
       color = ~League,
       colors = "Set1",
       text = ~Team) %>%
  add_text() %>%
  layout(title = "NMDS of the Baseball 2016 Data")
```

The plot shows no clear-cut pattern between the two leagues. However, the y-axis is better suited to differentiate between the leagues than the x-axis since more of the NL teams are on the negative side and vice versa for the AL teams. The outliers concerning this dimension are exclusively from the respective league. The most positive outliers are the AL teams Baltimore Orioles and Tamba Bay Rays. In the negative part the NL teams Atlanta Braves, San Francisco Giants, Arizona Biamondbacks, Miami Marlins and Colorado Rockies are disconnected from the other observations. Concerning the y-axis, the AL team Boston Red Sox are an outlier on the negative side and the NL teams Philadelphia Phillies, Milwaukee Brewers and San Diego Padres are on the positive end.

## 2.3)

Use Plotly to create a Shepard plot for the MDS performed and comment about how successful the MDS was. Which observation pairs were hard for the MDS to map successfully?

```{r shepard, message=FALSE}
shepard <- as.data.frame(Shepard(d, fit$points))
plot_ly(data = shepard,
        x = ~x,
        y = ~y,
        type = "scatter") %>%
  layout(title = "Shepard Plot",
         xaxis = list(title = "Original Minkowski Distance"),
         yaxis = list(title = "Distance after NMDS"))
```

The Shepard plot makes a quite good impression with the points giving the Minkowski distances and the distances after the NMDS constituing an overall monotonically increasing line over the whole range. However, there are some points closer to the origin of the plot that deviate negatively from the line. As can be seen by the final conversion value in the output of the NMDS, the final Stress value, is at about 15.9% which is considered an acceptable value.

## 2.4)

Produce series of scatterplots in which you plot the MDS variable that was the best in the differentiation between the leagues in step 2 against all other numerical variables of the data. Pick up two scatterplots that seem to show the strongest (positive or negative) connection between the variables and include them into your report. Find some information about these variables in Google – do they appear to be important in scoring the baseball teams? Provide some interpretation for the chosen MDS variable.

```{r scatters, message=FALSE, warning=FALSE}
OPS <- plot_ly(data = baseball,
               x = ~OPS,
               y = ~y,
               type = "scatter",
               name = ~"On Base + Slugging")
RBI <- plot_ly(data = baseball,
               x = ~RBI,
               y = ~y,
               type = "scatter",
              name = ~"Run Batted In")
subplot(OPS, RBI) %>%
  layout(title = "NMDS Y-Dimension and most important Correlates",
         yaxis = list(title = "NMDS Y-Dimension"))
```

I have selected two  of the variables with the highest correlations with the y-dimension of the presented NMDS. These are the sum of the percent of the time on base and the percent slugging (OPS, r = `r round(cor(baseball$OPS, baseball$x), digits = 2)`) and the runs a batter has made possible by his batting (RBI, r = `r round(cor(baseball$RBI, baseball$x), digits = 2)`). Both variables appear to be important statistics in baseball analysis and highly predictive of a teams success. For OPS, the ability of a player both to get on base and to hit for power, two important offensive skills, are represented. This measure is a composite measure of multiple key defining characteristics why its importance is not surprising. RBI credits a batter for making a play that allows a run to be scored.

# References

Ward, Grinstein and Keim 2015. "Interactive Data Visualization", London: Taylor & Francis, 2nd Edition, ISBN: 9781482257373

# Contributions

Assignment 1 was the primary responsibility of Varshith Konda while it was Malte Grönemanns on Assignment 2.