---
title: "Visualisation - Lab 2 - Group A7"
author: "Malte Grönemann and Varshith Konda"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(plotly)

olive <- read.csv("olive.csv")
```

# Assignment 1

## Introduction

The data contains the information about different contents of olive oils hailing from different regions in Italy, describing about different regions, areas and different acids present in them.

## 1.1

Create a scatterplot in Ggplot2 that shows dependence of Palmitic on Oleic in which
observations are colored by Linoleic. Create also a similar scatter plot in which you divide
Linoleic variable into fours classes (use cut_interval() ) and map the discretized variable to
color instead. How easy/difficult is it to analyze each of these plots? What kind of perception
problem is demonstrated by this experiment?



```{r 1.1, warning=FALSE}
graph1 <- ggplot(data = olive,
                 aes(x = palmitic,
                     y= oleic,
                     color=linoleic)) +
                geom_point() +
                labs(title = "1.1 | Dependency of palmitic on oleic",
                     x = "palmitic",
                     y = "oleic")
graph1
```



```{r 1.2, warning=FALSE}
    
Linoleic_interval_data <- cut_interval(olive$linoleic,4)
    
graph2 <- ggplot(data = olive,
                 aes(x = palmitic, 
                     y = oleic,
                     color = Linoleic_interval_data)) +
                  geom_point()+
  labs(title = "1.2 | Dependency of palmitic on oleic ",
                     x = "palmitic",
                     y = "oleic")
graph2
```



From  plot 1.1 where the points are colored by the Linoleic  we observe that due to the close
saturation of color ,it is difficult to differentiate between appropriate information from the plot. plot 1.2 is similar plot but colored by the dircretized varialble of Linoleic which indeed present us with more information regarding the distribution of the data.




## 1.2
Create scatterplots of Palmitic vs Oleic in which you map the discretized Linoleic with four
classes to:
a. Color
b. Size
c. Orientation angle (use geom_spoke() )
State in which plots it is more difficult to differentiate between the categories and
connect your findings to perception metrics (i.e. how many bits can be decoded by a
specific aesthetics

```{r 2.1, warning=FALSE}
graph3 <- ggplot(data = olive,
                 aes(x = palmitic,
                     y = oleic,
                     color = Linoleic_interval_data)) +
              geom_point()+
              labs(title= "1.3 | Palmitic and Oleic coloured by discretized lenoleic",
                   x="palmitic",
                   y="oleic")
graph3

graph4 <-  ggplot(data = olive,
                  aes(x = palmitic,
                      y = oleic,
                      size = Linoleic_interval_data)) +
                  geom_point()+
                   labs(title= "1.4 | Palmitic and Oleic size by discretized lenoleic",
                   x="palmitic",
                   y="oleic")
graph4

olive$linoleic_angle <- as.numeric(cut_interval(olive$linoleic,4,labels = c(0,.5*pi,1*pi, 2*pi)))


graph5 <-  ggplot(data = olive,
                  aes(x = palmitic,
                      y = oleic,
                      angle = linoleic_angle,
                      radius  = 50))+
          geom_point()+
                   labs(title= "1.5 | Palmitic and Oleic orientation angle by discretized lenoleic",
                   x="palmitic",
                   y="oleic")+
          geom_spoke()
graph5
```
Among the three plots plot 2.2 is hard to differentiate between the categories. due to the overlaping of data points and similar color for all classes.


## 1.3
Create a scatterplot of Oleic vs Eicosenoic in which color is defined by numeric values of
Region. What is wrong with such a plot? Now create a similar kind of plot in which Region is a
categorical variable. How quickly can you identify decision boundaries? Does preattentive or
attentive mechanism make it possible?

```{r 3.1, warning = FALSE}
graph6 <- ggplot(data = olive,
                 aes(x = oleic,
                     y = eicosenoic,
                     color = Region)) +
                     geom_point()+
          labs(title = "1.6 | Scatterplot of Oleic and Eicosenoic",
               x = "oleic",
               y= "eicosenoic")
graph6
```

## Converting numerical data into categorical data 

```{r 3.2 conversion}
region_categorial <- cut(olive$Region, 
                         breaks = c(0,1,2,3),
                         labels = c("North","South","Sardinia islands"))

graph7 <- ggplot(data = olive,
                 aes(x = oleic,
                     y = eicosenoic,
                     color = region_categorial)) +
              labs(title = "1.7 | Scatterplot of Oleic and Eicosenoic",
               x = "oleic",
               y= "eicosenoic")+
                  geom_point()
graph7
```




Decision boundaries can be easily defined in 1.7 to 1.6 since the 1.6 classified regions in a numerical value where it is difficult to interpret information from the graph compared to 3.2 where the regions are classified and color coded accordingly.

Both plots represent data in a similar way but  due to the pre-attentive mechanism,plot 1.7 makes things easier to identify on the data plot. 




## 1.4

Create a scatterplot of Oleic vs Eicosenoic in which color is defined by a discretized Linoleic (3
classes), shape is defined by a discretized Palmitic (3 classes) and size is defined by a
discretized Palmitoleic (3 classes). How difficult is it to differentiate between 27=3*3*3
different types of observations? What kind of perception problem is demonstrated by this
graph?

```{r 4.1}
linoleic_class_data <- cut_interval(olive$linoleic, 3,labels = c("low","medium","high"))
palmitic_class_data <- cut_interval(olive$palmitic, 3,labels = c("low","medium","high"))
palmitoleic_class_data <-as.numeric( cut_interval(olive$palmitoleic, 3,labels = c("low","medium","high")))

graph8 <- ggplot(data = olive,
                 aes(x = oleic,
                     y = eicosenoic,
                     color = linoleic_class_data,
                     shape = palmitic_class_data,
                     size = palmitoleic_class_data)) +
              labs(title ="1.8 | oleic vs eicosenoic",
                   x = "oleic",
                   y= "eicosenoic")+
                  geom_point()
graph8


```



The plot has over plotting which intern makes it difficult to see overlapping data points and it has too many classes to understand,due to overlaping issues perception of the data is confusing and this can lead wrong data interpretation.

The three classes should be plotted on different plot in order to get complete analysis of the data.

## 1.5

````{r question5}
graph9 <- ggplot(data = olive,
                 aes(x=oleic,
                     y= eicosenoic,
                     color =Region,
                     shape = palmitic_class_data,
                     size = as.numeric(palmitoleic_class_data))) +
  labs(title ="1.9 | oleic  vs eicosenoic",
        x = "oleic",
         y = "ecosenoic")+
                  geom_point()
graph9
``` 

Because of the variation in the data points color spectrum and cluster forming, and clear visual patterns as we can interpret the plot to some degree,
This approach is much better than plot 1.8 since we have less classes to grasp, region-based visual trends make this approach stronger
Even this plot has some issue of over plotting. 

from perspective of Treisman's theory our cognitive attention is limited by small amount of time ,where we can distinguish the change in color in similar sets of colors ,and accordingly we can easily identify the dark shaded region of the plot which has more color is accumulated.

## 1.6
 
```{r question 6, warning=FALSE}
graph10 <- plot_ly(data = olive, 
                   type='pie',
                   labels = ~Area,
                   values = olive,
                   showlegend = FALSE,
                   showlabels = FALSE,
                   textinfo = 'none'
                   )

graph10
```
As there are no labels and on the pie chart it is hard to understand the representation of the plot,As we cannot identify the angle differences with out labels this pie chart can be mislead portions of the chart,these plots are not ideal for static reports.


## 1.7

```{r question7}
graph11 <- ggplot(data = olive,
                  aes(x = linoleic,
                      y = eicosenoic)) +
                  geom_point() +
                  geom_density_2d()+
                  labs(title ="1.11 | 2d density plot of lenoleic and eicosenoic")

graph11
```

Contour lines does not describe particular relations of the plot points as we do not know the densities of the plot points ,when compared to the similar plot point we can see visual representation of regions divided into classes which gives more insight into the regional data than the 2d density plot.

# Assignment 2

## 2.1)

The data set baseball-2016.xlsx contains information about the scores of baseball teams in USA in 2016, such as: Games won, Games Lost, Runs peer game, At bats, Runs, Hits, Doubles, Triples, Home runs, Runs batted in, Bases stolen, Time caught stealing, Bases on Balls, Strikeouts, Hits/At Bats, On Base Percentage, Slugging percentage, On base+Slugging, Total bases, Double plays grounded into, Times hit by pitch, Sacrifice hits, Sacrifice flies, Intentional base on balls, and Runners Left On Base.

Load the file to R and answer whether it is reasonable to scale these data in order to perform
a multidimensional scaling (MDS).

```{r data2}
library(readxl)
baseball <- read_excel("baseball-2016.xlsx")
baseball_2 <- baseball[ , -(1:2)] # removing team name and league
baseball_2 <- scale(baseball_2)

anyNA(baseball_2)
```

The data is generally appropriate for multidimensional scaling (MDS) but some variables are count data. Therefore it is more appropriate to focus on order than distance which can be achieved by using non-metric MDS. Before doing the procedure, we want to exclude the variables defining the teams, the team name and the league they play in. Fortunately, the dataset contains no missing values that could mess with the procedure.

## 2.2)

Write an R code that performs a non-metric MDS with Minkowski distance=2 of the data (numerical columns) into two dimensions. Visualize the resulting observations in Plotly as a scatter plot in which observations are colored by League. Does it seem to exist a difference between the leagues according to the plot? Which of the MDS components seem to provide the best differentiation between the Leagues? Which baseball teams seem to be outliers?

```{r nmds, message=FALSE, warning=FALSE}
library(MASS)
library(plotly)

d <- dist(baseball_2, 
          method = "minkowski", 
          p = 2) # power of the minkowski distance set to 2

fit <- isoMDS(d, k = 2) # fitting a nmds for k = 2 dimensions

baseball$x <- fit$points[ , 1]
baseball$y <- fit$points[ , 2]

plot_ly(data = baseball,
       x = ~x,
       y = ~y,
       color = ~League,
       colors = "Set1",
       text = ~Team) %>%
  add_text() %>%
  layout(title = "NMDS of the Baseball 2016 Data")
```

The plot shows no clear-cut pattern between the two leagues. However, the y-axis is better suited to differentiate between the leagues than the x-axis since more of the NL teams are on the negative side and vice versa for the AL teams. The outliers concerning this dimension are exclusively from the respective league. The most positive outliers are the AL teams Baltimore Orioles and Tamba Bay Rays. In the negative part the NL teams Atlanta Braves, San Francisco Giants, Arizona Biamondbacks, Miami Marlins and Colorado Rockies are disconnected from the other observations. Concerning the y-axis, the AL team Boston Red Sox are an outlier on the negative side and the NL teams Philadelphia Phillies, Milwaukee Brewers and San Diego Padres are on the positive end.

## 2.3)

Use Plotly to create a Shepard plot for the MDS performed and comment about how successful the MDS was. Which observation pairs were hard for the MDS to map successfully?

```{r shepard, message=FALSE}
shepard <- as.data.frame(Shepard(d, fit$points))
plot_ly(data = shepard,
        x = ~x,
        y = ~y,
        type = "scatter") %>%
  layout(title = "Shepard Plot",
         xaxis = list(title = "Original Minkowski Distance"),
         yaxis = list(title = "Distance after NMDS"))
```

The Shepard plot makes a quite good impression with the points giving the Minkowski distances and the distances after the NMDS constituing an overall monotonically increasing line over the whole range. However, there are some points closer to the origin of the plot that deviate negatively from the line. As can be seen by the final conversion value in the output of the NMDS, the final Stress value, is at about 15.9% which is considered an acceptable value.

## 2.4)

Produce series of scatterplots in which you plot the MDS variable that was the best in the differentiation between the leagues in step 2 against all other numerical variables of the data. Pick up two scatterplots that seem to show the strongest (positive or negative) connection between the variables and include them into your report. Find some information about these variables in Google – do they appear to be important in scoring the baseball teams? Provide some interpretation for the chosen MDS variable.

```{r scatters, message=FALSE, warning=FALSE}
OPS <- plot_ly(data = baseball,
               x = ~OPS,
               y = ~y,
               type = "scatter",
               name = ~"On Base + Slugging")
RBI <- plot_ly(data = baseball,
               x = ~RBI,
               y = ~y,
               type = "scatter",
              name = ~"Run Batted In")
subplot(OPS, RBI) %>%
  layout(title = "NMDS Y-Dimension and most important Correlates",
         yaxis = list(title = "NMDS Y-Dimension"))
```

I have selected two  of the variables with the highest correlations with the y-dimension of the presented NMDS. These are the sum of the percent of the time on base and the percent slugging (OPS, r = `r round(cor(baseball$OPS, baseball$x), digits = 2)`) and the runs a batter has made possible by his batting (RBI, r = `r round(cor(baseball$RBI, baseball$x), digits = 2)`). Both variables appear to be important statistics in baseball analysis and highly predictive of a teams success. For OPS, the ability of a player both to get on base and to hit for power, two important offensive skills, are represented. This measure is a composite measure of multiple key defining characteristics why its importance is not surprising. RBI credits a batter for making a play that allows a run to be scored.

# Contributions

Assignment 1 was the primary responsibility of Varshith Konda while it was Malte Grönemanns on Assignment 2.