---
title: "Lab_3 Assignment 1"
author: "konda varshith , malte groenemann"
date: "21/09/2020"
output: html_document
runtime: shiny
---


```{r loading libraries , warnings = FALSE}
library(dplyr)
library(plotly)
library(ggplot2)
library(sf)
library(maps)
library(mapproj)



secrettoken = "pk.eyJ1IjoidmFya28xMTYiLCJhIjoiY2tmNWg2YjVqMDFsYTJzcjRqY2NxM2wyZiJ9.i_6sbQHwO3Sdsojg3TIdxw"
Sys.setenv('MAPBOX_TOKEN' = secrettoken)
aa_data <- read.csv("aegypti_albopictus.csv")

```


### Question 1 

Use MapBox interface in Plotly to create two dot maps (for years 2004 and 2013) that show
the distribution of the two types of mosquitos in the world (use color to distinguish between
mosquitos). Analyze which countries and which regions in these countries had high density of
each mosquito type and how the situation changed between these time points. What
perception problems can be found in these plots?
```{r 1.1 , warning = FALSE}



  data_2004 <- filter(aa_data,YEAR == 2004)


    plot1 <- map_data("world")%>%
  
           plot_mapbox(x= ~long,y=~lat)%>%
    
           add_markers(data = data_2004,x=~X,y=~Y,color=~VECTOR)%>%
      
           layout(title = "Plot - 1  mapping mosquitoes occurances in 2004 throughout the world")
    
           plot1




  data_2013 <- filter(aa_data,YEAR == 2013)

    plot2 <- map_data("world")%>%
          plot_mapbox(x =~lat , y= ~long)%>%
        
          add_markers(data = data_2013,x = ~X, y=~Y,color = ~VECTOR)%>%
      
          layout(title = "Plot - 2  mapping mosquitoes occurances in 2013 throughout the world")
  
          plot2


```

We can observe from the plot 1 that in 2004 the occurrences of the both mosquitoes spread throughout the continents ,Aedes aegypti has been occurred highly in coastal areas especially in the coast of southern america and countries of asia,kenya but on the 
contrary Taiwan and  had high occurrences of aedes abopictus.

aedes albopictus were mostly found in the more interior regions  and mostly found in  north america and coastal regions of japan and indonesia.


In plot 2 we can look at the data from the year 2013 where we can clearly see that the there was an exponential growth of aedes aegypti population in Brazil when compared to the population in  2004.Europe had seen increased number of aedes albopictus, overall there was decrease in aedes albopictus all over the world but taiwan had significantly same occurrences.

Both types of mosquitoes were found in central region of the globe and there were some occurrences in australia and europe.



# Question 2


Compute Z as the numbers of mosquitos per country detected during all study period. Use
plot_geo() function to create a choropleth map that shows Z values. This map should have an
Equirectangular projection. Why do you think there is so little information in the map?


```{r 2 , warning = FALSE}
compute_z <-   aa_data%>%count(COUNTRY_ID,sort = TRUE)


geo <- list(
  projection = list(
    type = 'equirectangular'
  ),
  showland = TRUE,
  
)




plot3 <- plot_geo(compute_z,type= 'choropleth')%>%
              add_trace(z = compute_z$n, locations = compute_z$COUNTRY_ID)%>%
              layout(geo = geo,title = "plot - 3  Mapping Z values on the map with equirectangualr projection")%>%
            colorbar(title = "Occurance of mosquitoes occurences in country")
  

plot3



```


In the plot 3 we mapped the Z values on the equirectangular projection of the world map , we can see that the variance in the indistinguishable color spectrum of the countries having less than five thousand occurrences is very low. this leads to perception issue in analyzing the plot.

Taiwan has seen high population of mosquitoes but the due to low contrast in map we cannot observe it with pre attentive cognitive  



# Question 3

Create the same kind of maps as in step 2 but use

a.)   Equirectangular projection with choropleth color log (𝑍)

b.)   Conic equal area projection with choropleth color log (𝑍)

Analyze the map from step 3a and make conclusions. Compare maps from 3a and 3b and
comment which advantages and disadvantages you may see with both types of maps.

```{r ,warning = FALSE}
## question - 3 

geo2 <- list(
  projection = list(
    type = 'equirectangular'
  ),
  showland = TRUE,
  landcolor = toRGB("gray95"),
  countrycolor = toRGB("gray80")
)


plot4 <- plot_geo(aa_data, type = 'choropleth')%>%
          add_trace(z= log(compute_z$n) , locations = compute_z$COUNTRY_ID )%>%
          layout(geo = geo2,title = "plot - 4  Mapping Log of Z values on the map with equirectangualr projection")
          
plot4  



geo3 <- list(
  projection = list(
    type = "conic equal area"
    
  ),
  showland = TRUE,
  landcolor = toRGB("gray95"),
  countrycolor = toRGB("gray80")
)


plot5 <- plot_geo(aa_data, type = 'choropleth')%>%
  add_trace(z= log(compute_z$n) , locations = compute_z$COUNTRY_ID )%>%
   layout(geo = geo3, title = "plot - 5  Mapping Z values on the map with conic equal area projection")
plot5 

```

In the plot 4  we mapped the Log(Z) values on the equirectangular projection of the world map  instead of computed values , using this approach it is easy to distinguish and analyze.

as the mapped regions have significant difference , it is easy to differentiate the intensity of mosquitoes occurrences on the map.

In plot 4 we can clearly see the entire map and analyze it easily as compared to plot 5 as it mapped with conic equal projections,
due to this we can only view certain section of the map, which is stressful task to get a clear perception of whole section of map at once, and it is difficult.




# Question 4

In order to resolve problems detected in step 1, use data from 2013 only for Brazil and

a.) Create variable X1 by cutting X into 100 piecies (use cut_interval() )

b.) Create variable Y1 by cutting Y into 100 piecies (use cut_interval() )

c.) Compute mean values of X and Y per group (X1,Y1) and the amount of observations N per group (X1,Y1)

d.)  Visualize mean X,Y and N by using MapBox

Identify regions in Brazil that are most infected by mosquitoes. Did such discretization
help in analyzing the distribution of mosquitoes?

```{r , warning = FALSE}

brazil_2013 <- filter(aa_data,aa_data$YEAR == 2013 & aa_data$COUNTRY == "Brazil")


brazil_2013$X1 <- cut_interval(brazil_2013$X,100)
brazil_2013$Y1 <- cut_interval(brazil_2013$Y,100)

brazil_2013_mean <- brazil_2013 %>%
  group_by(X1,Y1) %>%
  summarise(meanX = mean(X),meanY = mean(Y),
            N = n())

plot6 <- map_data("world")%>%
  plot_mapbox(x= ~long,y=~lat)%>%
  add_markers(data = brazil_2013_mean,x=~meanX,y=~meanY,color=~N)%>%
  layout(title= "plot - 5  Mapping Z values on the map with equirectangualr projection")

plot6
```