---
title: "Visualisation - Lab 3 - Group A7"
author: "Malte Grönemann and Varshith Konda"
date: "22/09/2020"
output: html_notebook
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(plotly)
library(sp)
library(sf)
library(maps)
library(mapproj)
```

# Assignment 1

File aegypti_albopictus.csv shows information about the location and detection time of two types of
mosquitoes. Both Aedes aegypti and Aedes albopictus mosquitoes may spread viruses like Zika,
dengue, chikungunya and other viruses but Aedes aegypti are more likely to spread these viruses
(and therefore are more dangerous).

```{r data1}
secrettoken = "pk.eyJ1IjoidmFya28xMTYiLCJhIjoiY2tmNWg2YjVqMDFsYTJzcjRqY2NxM2wyZiJ9.i_6sbQHwO3Sdsojg3TIdxw"
Sys.setenv('MAPBOX_TOKEN' = secrettoken)
aa_data <- read.csv("aegypti_albopictus.csv")
```

## 1.1

Use MapBox interface in Plotly to create two dot maps (for years 2004 and 2013) that show
the distribution of the two types of mosquitos in the world (use color to distinguish between
mosquitos). Analyze which countries and which regions in these countries had high density of
each mosquito type and how the situation changed between these time points. What
perception problems can be found in these plots?

```{r 1.1 , warning = FALSE}
data_2004 <- filter(aa_data,YEAR == 2004)

plot1 <- map_data("world")%>%
           plot_mapbox(x= ~long,y=~lat)%>%
           add_markers(data = data_2004,x=~X,y=~Y,color=~VECTOR)%>%
           layout(title = "Plot - 1  mapping mosquitoes occurances in 2004 throughout the world")
    
plot1


data_2013 <- filter(aa_data,YEAR == 2013)

plot2 <- map_data("world")%>%
          plot_mapbox(x =~lat , y= ~long)%>%
          add_markers(data = data_2013,x = ~X, y=~Y,color = ~VECTOR)%>%
          layout(title = "Plot - 2  mapping mosquitoes occurances in 2013 throughout the world")
  
plot2
```

We can observe from the plot 1 that in 2004 the occurrences of the both mosquitoes spread throughout the continents, Aedes aegypti has been occurring highly in coastal areas especially in the coast of South America and countries of Asia, Kenya but on the 
contrary Taiwan had high occurrences of aedes abopictus.

Aedes albopictus were mostly found in the more interior regions  and mostly found in  North America and coastal regions of Japan and Indonesia.

In plot 2 we can look at the data from the year 2013 where we can clearly see that the there was an exponential growth of aedes aegypti population in Brazil when compared to the population in  2004. Europe had seen increased number of aedes albopictus, overall there was decrease in aedes albopictus all over the world but Taiwan had significantly same occurrences.

Both types of mosquitoes were found in central region of the globe and there were some occurrences in Australia and Europe.

## 1.2

Compute Z as the numbers of mosquitos per country detected during all study period. Use
plot_geo() function to create a choropleth map that shows Z values. This map should have an
Equirectangular projection. Why do you think there is so little information in the map?


```{r 2 , warning = FALSE}
compute_z <-   aa_data%>%count(COUNTRY_ID,sort = TRUE)

geo <- list(
  projection = list(
    type = 'equirectangular'),
  showland = TRUE)

plot3 <- plot_geo(compute_z,type= 'choropleth')%>%
              add_trace(z = compute_z$n, locations = compute_z$COUNTRY_ID)%>%
              layout(geo = geo,
                     title = "plot - 3  Mapping Z values on the map with equirectangular projection")%>%
            colorbar(title = "Occurance of mosquitoes occurences in country")

plot3
```

In the plot 3 we mapped the Z values on the equirectangular projection of the world map, we can see that the variance in the indistinguishable color spectrum of the countries having less than five thousand occurrences is very low. This leads to perception issue in analyzing the plot.

Taiwan has seen high population of mosquitoes but the due to low contrast in map we cannot observe it with preattentive cognition.

## 1.3

Create the same kind of maps as in step 2 but use

a.)   Equirectangular projection with choropleth color log (𝑍)

b.)   Conic equal area projection with choropleth color log (𝑍)

Analyze the map from step 3a and make conclusions. Compare maps from 3a and 3b and
comment which advantages and disadvantages you may see with both types of maps.

```{r ,warning = FALSE}
geo2 <- list(
  projection = list(
    type = 'equirectangular'),
  showland = TRUE,
  landcolor = toRGB("gray95"),
  countrycolor = toRGB("gray80"))


plot4 <- plot_geo(aa_data, type = 'choropleth')%>%
          add_trace(z= log(compute_z$n) , 
                    locations = compute_z$COUNTRY_ID )%>%
          layout(geo = geo2,
                 title = "plot - 4  Mapping Log of Z values on the map with equirectangular projection")
          
plot4  


geo3 <- list(
  projection = list(
    type = "conic equal area"),
  showland = TRUE,
  landcolor = toRGB("gray95"),
  countrycolor = toRGB("gray80"))

plot5 <- plot_geo(aa_data, 
                  type = 'choropleth')%>%
  add_trace(z= log(compute_z$n) , 
            locations = compute_z$COUNTRY_ID )%>%
   layout(geo = geo3, 
          title = "plot - 5  Mapping Z values on the map with conic equal area projection")
plot5 

```

In the plot 4  we mapped the Log(Z) values on the equirectangular projection of the world map  instead of computed values , using this approach it is easy to distinguish and analyze.

as the mapped regions have significant difference , it is easy to differentiate the intensity of mosquitoes occurrences on the map.

In plot 4 we can clearly see the entire map and analyze it easily as compared to plot 5 as it mapped with conic equal projections,
due to this we can only view certain section of the map, which is stressful task to get a clear perception of whole section of map at once, and it is difficult.

Usage of this projection is unusual and it not necessary to have equal areas for the choropleth maps,therefore equirectangular projection is preffered.

## 1.4

In order to resolve problems detected in step 1, use data from 2013 only for Brazil and

a.) Create variable X1 by cutting X into 100 piecies (use cut_interval() )

b.) Create variable Y1 by cutting Y into 100 piecies (use cut_interval() )

c.) Compute mean values of X and Y per group (X1,Y1) and the amount of observations N per group (X1,Y1)

d.)  Visualize mean X,Y and N by using MapBox

Identify regions in Brazil that are most infected by mosquitoes. Did such discretization
help in analyzing the distribution of mosquitoes?

```{r , warning = FALSE}
brazil_2013 <- filter(aa_data,aa_data$YEAR == 2013 & aa_data$COUNTRY == "Brazil")

brazil_2013$X1 <- cut_interval(brazil_2013$X,100)
brazil_2013$Y1 <- cut_interval(brazil_2013$Y,100)

brazil_2013_mean <- brazil_2013 %>%
  group_by(X1,Y1) %>%
  summarise(meanX = mean(X),
            meanY = mean(Y),
            N = n())

plot6 <- map_data("world")%>%
  plot_mapbox(x= ~long,y=~lat)%>%
  add_markers(data = brazil_2013_mean,
              x=~meanX,
              y=~meanY,
              color=~N)%>%
  layout(title= "plot - 5  Mapping Z values on the map with equirectangualr projection")

plot6
```

We obsere high concentration of the points and points with brighest color are in coastal regions that are also highly populated , the dicretezisation reduced the over plotting when compared to plot 1 ,but improvement is only marginal , but the basic finding of the mosquitoe occurances are visible in  the plot 1 already. 


# Assignment 2

In this assignment, you will analyze the mean incomes of the Swedish households. Go to
http://www.scb.se and choose “English” language, and in the “Search” menu type “Disposable
income for households by region, type of households and age”, click “Search” and then click at
“Statistical Database”. Select “Mean value, SEK thousands”, all counties, age groups 18-29, 30-49 and
50-64, and year 2016. Download the “Comma delimited without heading” file.

## 2.1

Download a relevant map of Swedish counties from http://gadm.org/country and load it into
R. Read your data into R and process it in such a way that different age groups are shown in
different columns. Let’s call these groups Young, Adult and Senior.

```{r data2}
SCB_Incomes_2016 <- read.csv("SCB_Incomes_2016.csv", stringsAsFactors = FALSE)
SCB_Incomes_2016 <- spread(SCB_Incomes_2016, age, X2016)
colnames(SCB_Incomes_2016) <- c("Region", "Young", "Adult", "Senior")
Encoding(SCB_Incomes_2016$Region) <- "latin1"

SCB_Incomes_2016$Region <- str_remove(SCB_Incomes_2016$Region, " county")
SCB_Incomes_2016$Region <- str_remove(SCB_Incomes_2016$Region, "[:digit:][:digit:] ")
SCB_Incomes_2016 <- SCB_Incomes_2016[order(SCB_Incomes_2016$Region), ]

SE <- readRDS("gadm36_SWE_1_sf.rds")
SE$NAME_1[11] <- "Örebro"
```

## 2.2

Create a plot containing three violin plots showing mean income distributions per age group.
Analyze this plot and interpret your analysis in terms of income.

```{r violin}
ggplot(SCB_Incomes_2016) +
  geom_violin(aes(x = "Young", y = Young)) +
  geom_boxplot(aes(x = "Young", y = Young), width = .2) +
  geom_violin(aes(x = "Adult", y = Adult)) +
  geom_boxplot(aes(x = "Adult", y = Adult), width = .2) +
  geom_violin(aes(x = "Senior", y = Senior)) +
  geom_boxplot(aes(x = "Senior", y = Senior), width = .2) +
  labs(title = "Mean Income in Swedish Counties by Age Groups in 2016",
       x = "Age Category",
       y = "Mean Disposable HH Income (1000 SEK)") +
  xlim("Young", "Adult", "Senior") +
  theme_bw()
```

The violin plot shows that the Young earn considerably less than the other age groups. Even the lowest mean income household by county with heads in the upper two age groups have considerably higher incomes than the highest income county in the youngest category. Also the range of incomes is higher in the two older caterories. Incomes of the oldest seem to be slightly higher overall compared to the intermediate group. For all age groups, the typical distributional form of income can be observed: most density is concentrated in the lower part of the range while there are outliers at the top. The distributions are therefore positively skewed. This suggests considerable mean differences in income between counties.

## 2.3

Create a surface plot showing dependence of Senior incomes on Adult and Young incomes in
various counties. What kind of trend can you see and how can this be interpreted? Do you
think that linear regression would be suitable to model this dependence?

```{r surface}
plot_ly(z = as.matrix(SCB_Incomes_2016[2:4])) %>%
  add_surface()
```

Overall, we can see that the counties with higher incomes in the two lower age groups have higher earnings in the oldest age group as well. But there is a peak for high senior earnings at moderate earnings of the other groups. Using linear regression with income data is always a bit tricky because of the typical distributional form of income. This can be seen here as well, the relations seem not to be completely linear. When applied with some caution however, such relations could still be studied with a regression approach. For example with log-scaled income variables and non-linear specifications. 

## 2.4

Use Plotly and add_sf() function to visualize incomes of Young and Adults in two choropleth
maps. Analyze these maps and make conclusions. Is there any new information that you
could not discover in previous statistical plots?

```{r chloropleth}
SE$Young <- SCB_Incomes_2016$Young
SE$Adult <- SCB_Incomes_2016$Adult

Y <- plot_ly() %>%
  add_sf(data = SE,
         split = ~NAME_1,
         color = ~Young,
         showlegend = FALSE) %>%
  layout(title = list(text = "Mean Incomes of the Young Age Group"))

A <- plot_ly() %>%
  add_sf(data = SE,
         split = ~NAME_1,
         color = ~Adult,
         showlegend = FALSE) %>%
  layout(title = list(text = "Mean Incomes of the Adult Age Group"))

Y
A
```

It can now be seen that the outlier at the top in both age groups is Stockholm county. And in general, the mean income seems to be lower in the North and higher in the South. The second highest income county is Halland.

## 2.5

Use GPVisualizer http://www.gpsvisualizer.com/geocoder/ and extract the coordinates of
Linköping. Add a red dot to the choropleth map for Young from step 4 in order to show
where we are located.

```{r linköping}
lkpg <- list(latitude = 58.409814, longitude = 15.624522, name = "Linköping")

Y %>%
  add_markers(data = lkpg,
              x = ~longitude, 
              y = ~latitude,
              text = ~name,
              color = "red")
```

# Contributions

The main responsibility for Assignment 1 obliged to Varshith Konda while for Assignment 2 it was the responsibility of Malte Grönemann.